---
title: "Coronavirus analysis"
author: "Carl Howe, RStudio"
date: "3/7/2020"
output: html_document
params:
  input_country: "US"
  input_state: "Massachusetts"
  percent_susceptible_population: 0.8
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(dplyr)
library(DT)
library(ggplot2)
library(lubridate)
library(tidyr)
library(readr)
library(scales)
library(plotly)
library(here)
library(glue)
library(readr)
library(stringr)


# Function to read the raw CSV files. The files are aggregated to the country
# level and then converted to long format

clean_jhd_to_long <- function(df) {
    df_str <- deparse(substitute(df))
    var_str <- substr(df_str, 1, str_length(df_str) - 4)
    
    df %>% group_by(`Country/Region`) %>%
        filter(`Country/Region` != "Cruise Ship") %>%
        select(-`Province/State`, -Lat, -Long) %>%
        mutate_at(vars(-group_cols()), sum) %>% 
        distinct() %>%
        ungroup() %>%
        rename(country = `Country/Region`) %>%
        pivot_longer(
            -country, 
            names_to = "date_str", 
            values_to = var_str
        ) %>%
        mutate(date = mdy(date_str)) %>%
        select(country, date, !! sym(var_str)) 
}

confirmed_raw <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")
deaths_raw <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")

raw_coronavirus <- clean_jhd_to_long(confirmed_raw) %>%
    full_join(clean_jhd_to_long(deaths_raw)) %>%
    arrange(country, date)

coronavirus <- raw_coronavirus %>% 
    pivot_longer(confirmed:deaths, values_to = "cases", names_to = "type")

theme_set(theme_minimal())

un_country_list <- read_csv(here::here("data/UNdata_Export_20200316.csv"))
names(un_country_list) <- c("country", "year", "variant", "population")
un_country_list %>% filter(year == 2019)

coronavirus_countries <- coronavirus %>% 
    select(country) %>% 
    distinct(country)  ## for selecting a country

excluded_countries <- coronavirus_countries %>% 
    anti_join(un_country_list, by = c("country" = "country"))

countries_needing_translation <-read_csv(here::here("excluded_countries_dictionary.csv"))

select_countries <- c("All", coronavirus_countries, recursive = TRUE, use.names = FALSE)
select_countries <- tibble(select_countries) %>% 
    filter(select_countries != "Taiwan*" & select_countries != "Cruise Ship")
names(select_countries) <- c("country")
select_countries <- select_countries %>% 
    arrange(country)
full_countries <- select_countries %>% 
    left_join(countries_needing_translation, by = c("country")) %>% 
    mutate(UN_country = ifelse(is.na(UN_country), country, UN_country)) %>% 
    left_join(un_country_list, by = c("UN_country" = "country")) %>% 
    filter(year == "2019") 

nytimes <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv")
nytimes_data <- nytimes %>% 
  select(-fips)
nytimes_us_cases <- nytimes_data %>% 
  pivot_longer(cases:deaths,  values_to = "cases", names_to = "type")

state_populations <- read_csv(here::here("data/us-state-populations.csv"))

nytimes_us_cases <- nytimes_us_cases %>% 
  left_join(state_populations, by = "state") %>% 
  filter(!is.na(population))
```

## World-wide coronavirus model

This is a visualization of the current infection state using the simplest possible Susceptible-Infection-Recovery model. Data comes from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) [Coronavirus repository](https://github.com/CSSEGISandData/COVID-19). Country population information comes from the United Nations.

```{r worldwide_raw_data}
cv_ww <- coronavirus %>%
  group_by(date, type) %>% 
  summarize(cases = sum(cases)) %>% 
  select(date,type,cases) %>% 
  arrange(date)
cv_ww <- cv_ww %>% 
  mutate(type = ifelse(type == "confirmed", "infected", type))

cc_ww_wide <- cv_ww %>% 
  pivot_wider(names_from= type, values_from=c(cases)) %>% 
  arrange(date)

g <- ggplot(cv_ww, aes(x = date, y = cases, color = type, group = type)) + 
  geom_line() +
  scale_y_continuous(labels = comma_format(accuracy = 1)) +
  labs(title = "Worldwide coronavirus infections",
       caption = "Data sources: Johns Hopkins University and Rami Krispin's coronavirus package",
       color = "", x = "", y = "")

ggplotly(g, tooltip = c("x", "y", "group"))
```

## US coronavirus data

Here are the US raw numbers.

```{r us, echo=FALSE}
cv_us <- coronavirus %>%
  filter(country == "US") %>% 
  group_by(date, type) %>%
  drop_na(cases) %>% 
  summarize(cases = sum(cases, na.rm = TRUE)) %>% 
  select(date,type,cases) %>% 
  arrange(date)
cv_us <- cv_us %>% 
  mutate(type = ifelse(type == "confirmed", "infected", type))
g <- ggplot(cv_us, aes(x = date, y = cases, color = type, group = type)) + 
  scale_y_continuous(labels = comma_format(accuracy = 1)) +
  geom_line() +
  labs(title = "US Coronavirus Daily Updates",
       caption = "Data sources: Johns Hopkins University and Rami Krispin's coronavirus package",
       x = "", y = "", color = "")
ggplotly(g, tooltip = c("x", "y", "group"))

```

## Massachusetts coronavirus data

Here are the US raw numbers.

```{r massachusetts, echo=FALSE}
cv_ma <- nytimes_us_cases %>%
  filter(state == "Massachusetts") %>% 
  group_by(date, type) %>%
  drop_na(cases) %>% 
  summarize(cases = sum(cases, na.rm = TRUE)) %>% 
  select(date,type,cases) %>% 
  arrange(date)
cv_ma <- cv_ma %>% 
  mutate(type = ifelse(type == "cases", "infected", type))
g <- ggplot(cv_ma, aes(x = date, y = cases, color = type, group = type)) + 
  scale_y_continuous(labels = comma_format(accuracy = 1)) +
  geom_line() +
  labs(title = "Massachusetts Coronavirus Daily Updates",
       caption = "Data sources: New York Times",
       x = "", y = "", color = "")
ggplotly(g, tooltip = c("x", "y", "group"))

```


## Coronovirus parameters

```{r parameter_setting}
total_susceptible_population <- 263300000
percent_susceptible_population <- 0.80
infection_duration <- 14
r0 <- 2.2
beta <- r0 / infection_duration
gamma <- 1 / infection_duration
death_rate <- .02
projection_duration <- 180 #in days

parameter_string <- glue("Susceptible population = {prettyNum(total_susceptible_population, big.mark = ',')}, R0 = {r0}, infection duration = {infection_duration}, death rate = {death_rate}")
```

Now we'll model infection rates. The latest figures I've heard are an infection rate (R naught) of `r r0` people per infected person over the period in which they are infectious. We also assume a death rate of about `r round(death_rate * 100, 1)` percent and recovery and becoming non-infectious in `r infection_duration` days.

```{r ww_model}

ww_model <- cv_us
ww_model <- ww_model %>% 
   mutate(type = ifelse(type == "death", "deaths", type))

## extend the date by project_duration days

ww_wide <- ww_model %>% 
  pivot_wider(names_from= type, values_from=c(cases)) %>% 
  arrange(date)

ww_date_range <- tibble(date =  ymd(seq(min(ww_wide$date), (max(ww_wide$date)), "days")))

historic_range <- ww_date_range %>% 
  left_join(ww_wide, by = "date")

historic_rows <- nrow(historic_range)

ww_projection_dates <- tibble(date =  ymd(seq(min(ww_wide$date), (max(ww_wide$date) + projection_duration), "days")))

ww_range <- ww_projection_dates %>% 
  left_join(ww_wide, by = "date")

ww_projection <- ww_range %>% 
  mutate(total_infections     = infected,
         currently_infectious = infected - deaths,
         new_infections       = 0,
         new_deaths           = 0,
         total_deaths         = deaths,
         new_recoveries       = 0,
         total_recoveries     = 0,
         susceptible          = total_susceptible_population,
         ds_dt                = 0,
         di_dt                = 0,
         dr_dt                = 0) %>% 
  replace_na(list(cases_infected       = 0,
                  cases_deaths         = 0,
                  cases_recovered      = 0,
                  cumsum_infected      = 0,
                  cumsum_deaths        = 0,
                  cumsum_recovered     = 0,
                  total_infections     = 0, 
                  currently_infectious = 0,
                  new_infections       = 0, 
                  new_deaths           = 0, 
                  total_deaths         = 0,
                  new_recoveries       = 0,
                  total_recoveries     = 0)) %>% 
  arrange(date)
```

```{r model_math}

## run_model: Function to compute the SIR model on existing data.
## takes the following arguments
##
## f               the data frame with the historical data
## historical_row  the last row of the historical data
##
## All other parameters come from the global variables that define the disease in the chunk
## coronavirus parameters

## this is the function that does all the work on an historical infection dataframe
## and creates a forecast data frame.

run_model <- function(f, population, duration, rnaught, drate, proj_duration) {
    
    ww_model <- f
    ww_model <- ww_model %>% 
        mutate(type = ifelse(type == "death", "deaths", type))
    
    ## extend the date by project_duration days
    
    ww_wide <- ww_model %>% 
        pivot_wider(names_from= type, values_from=c(cases)) %>% 
        arrange(date)
    
    ww_date_range <- tibble(date =  ymd(seq(min(ww_wide$date), (max(ww_wide$date)), "days")))
    
    historic_range <- ww_date_range %>% 
        left_join(ww_wide, by = "date")
    
    historic_rows <- nrow(historic_range)
    
    ww_projection_dates <- tibble(date =  ymd(seq(min(ww_wide$date), (max(ww_wide$date) + proj_duration), "days")))
    
    ww_range <- ww_projection_dates %>% 
        left_join(ww_wide, by = "date")
    
    ## OK, here's the model
    ## Every day, we get new infections that's the (number of infectious people - deaths) * r0 / recovery_rate.
    ## New deaths is just the death_rate times infected people.
    ## New recoveries is the number of infected people from 14 days ago - number of deaths
    ## New number of infected people is prior number of infected people + new infections - deaths - new recoveries
    ## first, let's initialize our projected columns with the actual data.
    fnew <- ww_range %>% 
        mutate(total_infections     = infected,
               currently_infectious = infected - deaths,
               new_infections       = 0,
               new_deaths           = 0,
               total_deaths         = deaths,
               new_recoveries       = 0,
               total_recoveries     = 0,
               susceptible          = population - infected - deaths,
               ds_dt                = 0,
               di_dt                = 0,
               dr_dt                = 0) %>% 
        replace_na(list(cases_infected       = 0,
                        cases_deaths         = 0,
                        cases_recovered      = 0,
                        cumsum_infected      = 0,
                        cumsum_deaths        = 0,
                        cumsum_recovered     = 0,
                        total_infections     = 0, 
                        currently_infectious = 0,
                        new_infections       = 0, 
                        new_deaths           = 0, 
                        total_deaths         = 0,
                        new_recoveries       = 0,
                        total_recoveries     = 0)) %>% 
        arrange(date)
    
    ## SIR model parameter calculation
    
    beta <- as.numeric(rnaught / duration)
    gamma <- as.numeric(1 / duration)
    
    daily_infection_probability <- rnaught / duration
    
    for (i in (historic_rows + 1):nrow(fnew)) {
        
        ## Let's calculate the SIR model derivatives
        
        fnew$ds_dt[i] <- round(-beta * fnew$currently_infectious[i - 1] * fnew$susceptible[i - 1] / population) - fnew$new_deaths[i - 1]
        fnew$di_dt[i] <- round(-fnew$ds_dt[i] - gamma * fnew$currently_infectious[i - 1])
        fnew$dr_dt[i] <- round(gamma * fnew$currently_infectious[i - 1])
        
        ## Now update the state with derivatives
        
        fnew$new_infections[i]       <-  ifelse (fnew$di_dt[i] < 0, 0, fnew$di_dt[i])
        fnew$currently_infectious[i] <-  fnew$currently_infectious[i - 1] + fnew$di_dt[i]
        fnew$new_deaths[i]           <-  round(fnew$currently_infectious[i - 1] * drate / duration)
        fnew$new_recoveries[i]       <-  fnew$dr_dt[i]
        fnew$susceptible[i]          <-  fnew$susceptible[i - 1]      + fnew$ds_dt[i]
        fnew$total_recoveries[i]     <-  fnew$total_recoveries[i - 1] + fnew$new_recoveries[i]
        fnew$total_deaths[i]         <-  fnew$total_deaths[i - 1]     + fnew$new_deaths[i]
        fnew$total_infections[i]     <-  fnew$total_infections[i - 1] + fnew$new_infections[i]
    }
    return(fnew)
}

derive_country_stats <- function() {     ## for debugging
  if (params$input_country == "All") {
    df <- coronavirus %>%
      group_by(date, type) %>% 
      summarize(cases = sum(cases, na.rm = TRUE)) %>% 
      select(date,type,cases) %>% 
      arrange(date)
  } else {
    df <- coronavirus %>%
      filter(country == params$input_country) %>% 
      group_by(date, type) %>%
      drop_na(cases) %>% 
      summarize(cases = sum(cases, na.rm = TRUE)) %>% 
      select(date,type,cases) %>% 
      arrange(date)
  }
  df <- df %>% 
    mutate(type = ifelse(type == "confirmed", "infected", type))
  df
}

model_builder <- function(){
  
  country_info <- full_countries %>% 
    filter(country == params$input_country) %>% 
    head(1)
  
  df <- derive_country_stats()
  run_model(df, total_susceptible_population, infection_duration,
            r0, death_rate, projection_duration)
}


```

```{r state_computations}

derive_state_stats <- function() {     ## for debugging
  if (params$input_state == "All") {
    df <- nytimes_us_cases %>%
      group_by(date, type) %>% 
      summarize(cases = sum(cases, na.rm = TRUE)) %>% 
      select(date,type,cases) %>% 
      arrange(date)
  } else {
    df <- nytimes_us_cases %>%
      filter(state == params$input_state) %>% 
      group_by(date, type) %>%
      drop_na(cases) %>% 
      summarize(cases = sum(cases, na.rm = TRUE)) %>% 
      select(date,type,cases) %>% 
      arrange(date)
  }
  df <- df %>% 
    mutate(type = ifelse(type == "cases", "infected", type))
  df
}

state_model_builder <- function(){
  
  state_info <- nytimes_us_cases %>% 
    filter(state == params$input_state) %>% 
    head(1)
  population <- state_info$population * percent_susceptible_population/100
  df <- derive_state_stats()
  run_model(df, population, infection_duration,
            r0, death_rate, projection_duration)
}

```


```{r plot_all_us}
options(scipen = 999)
model_result <- model_builder()

## write.csv(model_result, here::here("model-result.csv"))
final_values <- model_result %>% 
  filter(date == max(model_result$date)) %>% 
  head(n = 1)       ## in case we get multiple results

## name mapping

name_map <- tribble( ~column_name, ~`Model Output`,
                     "total_infections",      "Total Infections",
                     "currently_infectious",  "Currently Infectious",
                     "new_infections",        "New Infections",
                     "new_deaths",            "New Deaths",
                     "total_deaths",          "Total Deaths",
                     "new_recoveries",        "New Recoveries",
                     "total_recoveries",      "Total Recoveries",
                     "susceptible",           "Suspectible Population")

projection_long <- model_result %>% 
  pivot_longer(total_infections:total_recoveries)

max_infectious <- model_result %>% 
  filter(model_result$currently_infectious == max(model_result$currently_infectious)) %>% 
  head(n = 1)       ## in case we get multiple results

final_values <- model_result %>% 
  filter(date == max(model_result$date)) %>% 
  head(n = 1)       ## in case we get multiple results

max_plot_y_value <- max(projection_long$value) * 0.90
max_infection_text_x_value <- max_infectious$date - days(3)

country_info <- full_countries %>% 
  filter(country == params$input_country) %>% 
  head(1)
total_susceptible_population <- country_info$population * params$percent_susceptible_population * 1000

plot_df <- projection_long %>% 
  left_join(name_map, by = c("name" = "column_name"))

g <-ggplot(plot_df, aes(x = date, y = value, color = `Model Output`, group = `Model Output`)) +
  geom_line() +
  scale_y_continuous(labels = unit_format(accuracy = 0.1, unit = "M", scale = 1e-6)) +
  scale_x_date(date_breaks = "2 weeks", limits = c(min(projection_long$date), max(projection_long$date) + days(10)), date_labels = "%B %d",
               expand = expansion(0, 0.4)) +
  scale_color_brewer(palette = "Dark2") +
  annotate("segment", x = max_infectious$date, xend = max_infectious$date, y = 0, yend = max_plot_y_value * 0.9, color = "gray20",  linetype = "dotted") +
  annotate("text",  x = max_infection_text_x_value, y = max_plot_y_value, 
           color = "gray20", size = 2.75,  hjust = "right", vjust = "inward",
           label = paste0("Peak infectious population\n",
                          prettyNum(round(max_infectious$currently_infectious), big.mark = ","),
                          "\non ", format(max_infectious$date, "%B %d"))) +
  annotate("text", x = max_infection_text_x_value + days(30), y = final_values$total_deaths * 2.5, 
           color = "gray20", size = 2.75, hjust = "inward", vjust = "inward",
           label = paste0("Total deaths\n", prettyNum(round(final_values$total_deaths), big.mark = ","))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
  labs(title = paste0(params$input_country, " Coronavirus Forecast\n"),
       x = "", y = "", color = "")

ggplotly(g, tooltip = c("x", "y", "group"))
```


```{r plot_all_massachusetts}
options(scipen = 999)
model_result <- state_model_builder()

## write.csv(model_result, here::here("model-result.csv"))
final_values <- model_result %>% 
  filter(date == max(model_result$date)) %>% 
  head(n = 1)       ## in case we get multiple results

## name mapping

name_map <- tribble( ~column_name, ~`Model Output`,
                     "total_infections",      "Total Infections",
                     "currently_infectious",  "Currently Infectious",
                     "new_infections",        "New Infections",
                     "new_deaths",            "New Deaths",
                     "total_deaths",          "Total Deaths",
                     "new_recoveries",        "New Recoveries",
                     "total_recoveries",      "Total Recoveries",
                     "susceptible",           "Suspectible Population")

projection_long <- model_result %>% 
  pivot_longer(total_infections:total_recoveries)

max_infectious <- model_result %>% 
  filter(model_result$currently_infectious == max(model_result$currently_infectious)) %>% 
  head(n = 1)       ## in case we get multiple results

final_values <- model_result %>% 
  filter(date == max(model_result$date)) %>% 
  head(n = 1)       ## in case we get multiple results

max_plot_y_value <- max(projection_long$value) * 0.90
max_infection_text_x_value <- max_infectious$date - days(3)

state_info <- nytimes_us_cases %>% 
    filter(state == params$input_state) %>% 
    head(1)
total_susceptible_population <- state_info$population * params$percent_susceptible_population

plot_df <- projection_long %>% 
  left_join(name_map, by = c("name" = "column_name"))

g <-ggplot(plot_df, aes(x = date, y = value, color = `Model Output`, group = `Model Output`)) +
  geom_line() +
  scale_y_continuous(labels = unit_format(accuracy = 0.1, unit = "M", scale = 1e-6)) +
  scale_x_date(date_breaks = "2 weeks", limits = c(min(projection_long$date), 
                                                   max(projection_long$date) + days(10)), 
               date_labels = "%B %d",
               expand = expansion(0, 0.4)) +
  scale_color_brewer(palette = "Dark2") +
  annotate("segment", x = max_infectious$date, xend = max_infectious$date, 
           y = 0, yend = max_plot_y_value * 0.9, color = "gray20",  linetype = "dotted") +
  annotate("text",  x = max_infection_text_x_value, y = max_plot_y_value, 
           color = "gray20", size = 2.75,  hjust = "right", vjust = "inward",
           label = paste0("Peak infectious population in ", params$input_state, "\n",
                          prettyNum(round(max_infectious$currently_infectious), big.mark = ","),
                          "\non ", format(max_infectious$date, "%B %d"))) +
  annotate("text", x = max_infection_text_x_value + days(30), y = final_values$total_deaths * 2.5, 
           color = "gray20", size = 2.75, hjust = "inward", vjust = "inward",
           label = paste0("Total deaths\n", prettyNum(round(final_values$total_deaths), big.mark = ","))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
  labs(title = paste0(params$input_state, " Coronavirus Forecast\n"),
       x = "", y = "", color = "")

ggplotly(g, tooltip = c("x", "y", "group"))
```


```{r short_term}
short_duration <- 30
projection_long <- model_result %>% 
  pivot_longer(total_infections:total_recoveries) %>%
  filter(date <= date(now() + days(short_duration))) %>% 
  select(date, name, value)

plot_df <- projection_long %>% 
  left_join(name_map, by = c("name" = "column_name"))

g <-ggplot(plot_df, aes(x = date, y = value, color = `Model Output`, group = `Model Output`)) +
  geom_line() +
  scale_y_continuous(limits = c(0, max(projection_long$value)), labels = scales::comma) +
  scale_x_date(date_breaks = "1 week", limits = c(date(min(projection_long$date)), date(now() + days(short_duration))), date_labels = "%B %d") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
g <- g + labs(title = paste0(short_duration, "-Day US COVID-19 SIR Model Forecast"),
              caption = paste0(parameter_string, "\nData sources: Johns Hopkins University and Rami Krispin's coronavirus package"),
              x = "", y = "", color = "")
ggplotly(g, tooltip = c("x", "y", "group"))

```

```{r short_term_2}

todays_values <- derive_country_stats() %>% 
  ungroup() %>% 
  filter(date == max(date)) %>% 
  mutate(type = ifelse(type == "death", "Deaths", str_to_title(type)),
         casestring = prettyNum(abs(cases), big.mark = ","),
         datestring = format(date, "%B %d"))

todays_values_x_value <- todays_values$date - days(3)
todays_values_string <-  paste0("Reported on ", first(todays_values$datestring), "\n", 
                                paste(with(todays_values, glue("{type}: {casestring}\n")), collapse="\n"))
model_result <- model_builder() %>% 
  filter(date <= todays_values_x_value + days(33))

final_values <- model_result %>% 
  filter(date == max(model_result$date)) %>% 
  head(n = 1)       ## in case we get multiple results

## name mapping

name_map <- tribble( ~column_name, ~Category,
                     "total_infections",      "Total Infections",
                     "currently_infectious",  "Currently Infectious",
                     "new_infections",        "New Infections",
                     "new_deaths",            "New Deaths",
                     "total_deaths",          "Total Deaths",
                     "new_recoveries",        "New Recoveries",
                     "total_recoveries",      "Total Recoveries",
                     "susceptible",           "Suspectible Population")

projection_long <- model_result %>% 
  pivot_longer(total_infections:total_recoveries)

max_infectious <- model_result %>% 
  filter(model_result$currently_infectious == max(model_result$currently_infectious)) %>% 
  head(n = 1)       ## in case we get multiple results

final_values <- model_result %>% 
  filter(date == max(model_result$date)) %>% 
  head(n = 1)       ## in case we get multiple results

max_plot_y_value <- max(projection_long$value) * 0.90
max_infection_text_x_value <- max_infectious$date - days(3)

country_info <- full_countries %>% 
  filter(country == params$input_country) %>% 
  head(1)

plot_df <- projection_long %>% 
  left_join(name_map, by = c("name" = "column_name")) %>% 
  select(date, value, Category)

g <-ggplot(plot_df, aes(x = date, y = value, color = Category, group = Category)) +
  geom_line() +
  scale_y_continuous(label = comma) +
  scale_x_date(date_breaks = "2 weeks", limits = c(min(projection_long$date), todays_values_x_value + days(33)), date_labels = "%B %d",
               expand = expansion(0, 0.4)) +
  scale_color_brewer(palette = "Dark2") +
  
  ## annotate for latest reported values
  
  annotate("segment", x = todays_values$date, xend = todays_values$date, y = 0, yend = max_plot_y_value * 0.5, color = "gray20",  linetype = "dotted") +
  annotate("text",  x = todays_values_x_value, y = max_plot_y_value * 0.65, 
           color = "gray30", size = 2.75,  hjust = "right", vjust = "inward",
           label = todays_values_string) +
  
  ## and for peak infectious population
  
  annotate("segment", x = max_infectious$date, xend = max_infectious$date, y = 0, yend = max_plot_y_value * 0.9, color = "gray20",  linetype = "dashed") +
  annotate("text",  x = max_infection_text_x_value - days(7), y = max_plot_y_value, 
           color = "gray20", size = 2.75,  hjust = "inward", vjust = "inward",
           label = paste0("On ",format(max_infectious$date, "%B %d"),
                          "\nInfected: ", prettyNum(round(max_infectious$currently_infectious), 
                                                    big.mark = ","),
                          "\nDeaths: ", prettyNum(round(max_infectious$total_deaths), big.mark = ","),
                          "\non ", format(max_infectious$date, "%B %d"))) +
  
  ## and for total deaths            
  
  annotate("text", x = max_infection_text_x_value + days(30), y = final_values$total_deaths * 3, 
           color = "gray20", size = 2.75, hjust = "inward", vjust = "inward",
           label = paste0("Total deaths\n", prettyNum(round(final_values$total_deaths), big.mark = ","))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "bottom") +
  labs(x = "", y = "", color = "")

ggplotly(g, tooltip = c("x", "y", "group"))

```

